Analysis of competitive surfing tournaments with generalized Bradley-Terry likelihoods

### 2022-04-24 inst/surfing_jsa_rebuttal_referee2.tex

Section 7, while an interesting topic, feels like it falls more into
the realm of speculation than concrete analysis in the way it is
presented. The methodology is not very clear and it would be nice to
see a more explicit definition and derivation of pc

\textcolor{blue}{We agree that this section is more speculative than
  the remainder of the manuscript.  However, we would reiterate that
  the topic is clearly interesting and worth investigating, as both
  referees note.  Non-competitive behaviour of the sort discussed here
  is notoriously difficult to study; and in particular, formal
  statistical analysis of non-competitive behaviour is almost
  non-existant in the literature.  We are not aware of any other
  probability model that is able to even establish the existence of
  non-competitive behaviour, let alone quantifiy it.  We believe that
  our approach is new, and our probability model for non-competitive
  behaviour appropriate and plausible.\\ But, on re-reading the
  section, we agree that the presentation is indeed somewhat obscure.
  We have extensively re-written the section and believe it is much
  improved.  We include a new table motivating and discussing
  non-competitive behaviour using a particular example from the
  tournament.  We note that we presented a likelihood function for the
  three-competitor case in the first draft (it is now Table 3), and we
  consider this to be as clear as reasonably practicable.}

Overall, it feels like the main conclusion is ``not all surfers have
the same skill". This is not insignificant, however, it is not
particularly surprising and is one that could have been reached using
a number of different methods. Had the authors done a more complete
job of fleshing out why the Bradley-Terry approach was better for
reaching this conclusion than others, I would be more satisfied, but
as it stands it falls a little flat.

\textcolor{blue}{We regard the novelty of our computational approach
  as being access to the full likelihood function over allowable
  Bradley-Terry strengths (that is, not restricted to the MLE or point
  hypotheses, but over the entire simplex, $\left\lbrace
  p_1,\ldots,p_n\left|p_i\geq 0,\sum p_j=1\right.\right\rbrace$) and
  also the furnishing of compact and natural R idiom to allow the
  creation of such likelihood functions.  As such, there are many many
  statistical tests that could be conceived and executed using our
  method.  Our original priority was to submit a brief and compact
  manuscript but on reflection the referee seems to be saying that we
  have erred on the side of brevity.  We have added a topical
  statistical test, assessing Medina's performance at Lemoore as an
  example of the power of our method.  This test returned a
  significant p-value, a result not achievable with other methods.  We
  feel that this new analysis complements the existing analysis
  comparing Medina with Ferreira (and which also rejected the null);
  we have added a short discussion to the manuscript.  \\ \\ Reading
  between the lines of the comment, we detect a certain disappointment
  that some of the results were non-significant.  We too felt a little
  chagrined that our method returned a non-significant result for the
  compatriot analysis.  However, we would argue that it is important
  to report all findings, significant or not---especially for
  difficult or otherwise challenging hypotheses not accessible to
  conventional techniques.  We would also point out that even for
  nonsiginficant results, our methods were considerably more
  informative than ``failure to reject a null'': as we stated in the
  submission, we present a novel quantification of noncompetitiveness
  (viz, $p_C$), and further, presented a nonzero maximum likelihood
  estimate for $p_C$ which, we suggest, is a worthy and interesting
  finding.  We also point out that such noncompetitiveness is a
  general phenomenon in the more general context of tournaments.  We
  present our work as the first application of a novel and very
  general method: we hope that subsequent analysis would cite this
  work.}

### 2022-04-24 inst/surfing_jsa.tex
more rebuttal

Medina's performance was startlingly good at Freshwater Pro ({\tt
  wct08}).  Indeed, Medina came first at only two venues (the other
was at Jeffries Bay, South Africa) and it is natural to enquire
whether there is statistical evidence to support the assertion that he
performed better under the more controlled conditions of the Lemoore
Surf Ranch.  To do this, we create an artificial surfer, {\tt
  Medina\_wct08}, who competes at Lemoore, and test the null that {\tt
  Medina} has the same Plackett-Luce strength as {\tt Medina\_wct08}.
We reject this null at a $p$-value of 0.043 and conclude that Medina's
performance at Lemoore is significantly better than his performance at
natural venues.

>>>>

Medina's performance was startlingly good at Freshwater Pro ({\tt
  wct08}).  Indeed, Medina came first at only two venues (the other
was at Jeffries Bay, South Africa) and it is natural to enquire
whether there is statistical evidence to support the assertion that he
performed better under the more controlled conditions of the Lemoore
Surf Ranch.  To do this, we create an artificial surfer, {\tt
  Medina\_wct08}, who competes at Lemoore, and test the null that {\tt
  Medina} has the same competitive strength as {\tt Medina\_wct08}.
We reject this null at a $p$-value of 0.043 and conclude that Medina's
performance at Lemoore is significantly better than his performance at
natural venues.  From an inferential perspective, this test is
interesting because competitor {\tt Medina} never faces {\tt
  Medina\_wct08}.  It is as though competitor A consistently beats B,
and B consistently beats C in pairwise competition, but A has never
directly competed against C.  Nevertheless we may assert with some
confidence that the Bradley-Terry strength of A exceeds that of C in
the absence of nontransitive effects~\citep{west2008}, via indirect
comparison.
